The \dealii Library, Version 9.7

### 2025-07-03 9.7/paper.tex
Add some of the new features of 9.7.

@@

Many of the \dealii tutorial programs were revised in a variety of ways
as part of this release: Around 130 of the more than 1600 (non-merge)
commits that went into this release touched the tutorial.
% data generated using these commands:
% - tutorial: git log --since 2023/07/07 --until 2024/08/11 --no-merges examples | grep commit | wc -l
% - total:    git log --since 2023/07/07 --until 2024/08/11 --no-merges | grep commit | wc -l
In addition, there are a number of new tutorial
programs:
\begin{itemize}
  \item
    \step{93}
    demonstrates how one can implement ...
    \todo{Complete}
    \step{93} was written by Sam Scheuerman.
  \item
    \step{97}
    is a program that solves ... curl-curl ...
    \todo{Complete}
    \step{93} was written by Siarhei Uzunbajakau.
\end{itemize}

@@

D.~Arndt and B.~Turcksin: Research sponsored by the Laboratory Directed Research and
Development Program of Oak Ridge National Laboratory, managed by UT-Battelle,
LLC, for the U. S. Department of Energy.

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Work towards C++20 modules}
\label{subsec:modules}

\dealii{} currently requires a compiler that understands C++17, but
some additional features are enabled if a compiler supports the C++20
standard. C++20 also introduces ``modules'', a way by which software
packages can export their public interface without relying on the
mechanism of textual inclusion of header files that C++ has inherited from C. The
end goal of C++20 modules is that one can replace the use of long
lists of statements such as\\
\hspace*{1cm}\texttt{\#include <deal.II/grid/tria.h}\\
by a simple\\
\hspace*{1cm}\texttt{import dealii;}

A lot of effort -- perhaps 6 weeks of full-time work, and around 150
pull requests -- has gone into evaluating whether \dealii{} can be
built using modules (in addition to header files, which will need to
be provided for a long time to come for backward compatibility). A
detailed accounting of this effort can be found in 
\cite{bangerth2025experienceconvertinglargemathematical}. The majority
of this work addressed features of the \dealii{} code base,
accumulated over its more than 25 year history, that were acceptable
in a header-based system but not in a module-based system. To name
just one of many examples, there can be cycles of header files that mutually
\texttt{\#include} each other, but this is not allowed for
modules. Most of the pull requests for this project therefore simply
cleaned up our code base, though without having to introduce any
incompatibilities.

As shown in \cite{bangerth2025experienceconvertinglargemathematical},
the use of C++20 modules
reduces compile times for the library itself, though the evidence is
mixed for tutorial programs or applications that build on
\dealii{}. Given that at the moment few systems have compilers, CMake
versions, and build systems that are sufficiently new to support
module builds, it will be several years before this feature will
become widely available. However, the infrastructure for it is now
largely in place.

@@

Many of the \dealii tutorial programs were revised in a variety of ways
as part of this release: Around 130 of the more than 1600 (non-merge)
commits that went into this release touched the tutorial.
% data generated using these commands:
% - tutorial: git log --since 2023/07/07 --until 2024/08/11 --no-merges examples | grep commit | wc -l
% - total:    git log --since 2023/07/07 --until 2024/08/11 --no-merges | grep commit | wc -l
In addition, there are a number of new tutorial
programs:
\begin{itemize}
  \item
    \step{93}
    demonstrates how one can implement problems in which some of the
    variables to be solved for are not degrees of freedom of a finite
    element field (so-called ``non-local'' degrees of freedom, because
    they have no associated node functional that is tied to a cell or
    a patch of cells). Specifically, the program solves an
    optimization problem in which some of the variables are scalars
    that multiply certain source terms.
    \step{93} was written by Sam Scheuerman.
  \item
    \step{97}
    is a program that solves a curl-curl problem from electromagnetics
    -- more specifically, from magnetostatics. The program shows a
    formulation in which two curl-curl problems are solved
    consecutively to model the magnetic field that results from the
    current in an electrically conducting coil.
    \step{97} was written by Siarhei Uzunbajakau.
\end{itemize}

@@

\todo[inline]{Everyone please update as appropriate.}
D.~Arndt and B.~Turcksin: Research sponsored by the Laboratory Directed Research and
Development Program of Oak Ridge National Laboratory, managed by UT-Battelle,
LLC, for the U. S. Department of Energy.
